---
title: "Take-home Exercise 2"
author: "Stefanie Felicia"
date: "May, 12 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## 1.1 Background

Over the past decade, the community of Oceanus has undergone transformations and challenges evolving from its fishing-centric origins. Currently, people are shifting to another investment, i.e. ocean tourism industry, which creates a growing tension. With the growing tourism, it attracted international pop star Sailor Shift who wanted to shoot his/her new music video in the island.

Clepper Jensen is a former analyst at FishEye company who is new a seasoned journalist for Hecklee Herald, is observing the tension and recently looking at temporary closure of Nemo Reef. By using investigative tools and radio communication, he uncovered a complex web of expedited approvals and secretive logistics, in which I will be helping him to develop visualization to help give better clarity to the story.

### 1.2 Objective

Objective of this exercise:

-   Use visual analytics to help Clepper understand and explore the interactions and relationships between people and vessels that relates to environmentalism (known associates of Green Guardians), Sailor Shift, and fishing/leisure vessels.
-   Nadia Conti was previously entangled in an illegal fishing scheme, it is suspected that she may have continued illegal activity within Oceanus, so we want to use visual analytics to provide evidence whether she's doing something illegal.

### 1.3 The Data

### 1.4 Load Required Libraries

```{r}
pacman::p_load(tidyverse, jsonlite, SmartEDA, tidygraph, ggraph, knitr, DT, stringr, cronologia, tidytext, dplyr)
```

The libraries used in this exercise are the following:

```{r, echo=FALSE}
table_data <- matrix(
  c("jsonlite",      "Convert JSON data to R objects",
    "tidyverse",  "Provide key data transformation functions",
    "SmartEDA",     "Help in getting the complete exploratory data analysis by running the function instead of writing long R codes", 
    "tidygraph",    "Provides tidy API for network or graph manipulation",
    "ggraph", "Extension of the ggplot2 API tailored to graph visualizations", 
    "knitr",  "Provides a tool for dynamic report generation",
    "datable", "Provides an R interface to the JavaScript library DataTables", 
    "stringr", "Provides a cohesive set of functions designed to make working with strings easy", 
    "cronologia", "HTML vertical timeline from a data frame as an input for 'rmarkdown' documents and 'shiny' applications", 
    "tidytext", "make many text mining tasks easier, more effective, and consistent with tools already in wide use",  
    "dplyr", " grammar of data manipulation, providing a consistent set of verbs that help solve the most common data manipulation challenges"), 
  nrow = 11,
  byrow = TRUE
)

kable(table_data, col.names = c("Package Name", "Description"))
```

### 1.5 Import Data

In this exercise, we'll be using the mc3.json file and extract data using the jsonlite library.

```{r}
MC3 <- fromJSON("data/MC3_graph.json")
MC3_schema <- fromJSON("data/MC3_schema.json")
```

## 2.1 Inspect Knowledge Graph Structure

Before preparing the data, we'll check the structure of the knowledge graph.

```{r}
glimpse(MC3)
```

Since the industry column is in list data type, we'' exclude the column since list data type isn't acceptable by `tbl_graph()`.

## 2.2 Extract Edges and Nodes Tables

The `as_tibble()` function is used to extract nodes and links tibble data frame from MC3 tibble data frame.

```{r}
mc3_nodes <- as_tibble(MC3$nodes)
mc3_edges <- as_tibble(MC3$edges)
```

## 2.3 Initial EDA

We'll use `ExpCatViz()` from SmartEDA library to reveal frequency distribution of all categorical fields from mc3_nodes.

```{r}
ExpCatViz(data=mc3_nodes,
          col="lightblue")
```

## 2.4 Data Cleaning and Wrangling

The following steps are taken to clean the **nodes**:

-   Convert values in id into character data type

-   Exclude record with id that are missing

-   Exclude records with similar id values

-   Exclude *thing_collected* field

```{r}
mc3_nodes_cleaned <- mc3_nodes %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(id)) %>%
  distinct(id, .keep_all = TRUE) %>%
  select(-thing_collected)
```

Next, the following steps are taken to clean the **edges**:

-   Rename source *column* into *from_id* and *target* column into *to_id*

-   Convert values in *from_id* and *to_id* column into character data type

-   Exclude values in from_id and to_id that's not found in the id column from *mc3_nodes_cleaned*

-   Exclude records where *from_id* and *to_id*, or *from_id* or *to_id* is missing

```{r}
    mc3_edges_cleaned <- mc3_edges %>%
      rename(from_id = source, 
             to_id = target) %>%
      mutate(across(c(from_id, to_id), 
                    as.character)) %>%
      filter(from_id %in% mc3_nodes_cleaned$id, 
             to_id %in% mc3_nodes_cleaned$id) %>%
      filter(!is.na(from_id), !is.na(to_id))
```

A new dataframe from the existing data frame (*mc3_nodes_cleaned*) and *.row_id* column is added which assigns unique row number to each row number (works like row index). Then we select *id* and .*row_id* columns whereas other columns are dropped.

```{r}
node_index_lookup <- mc3_nodes_cleaned %>%
  mutate(.row_id = row_number()) %>%
  select(id, .row_id)
```

Then from_id and to_id columns in mc_edges_indexed are converted it to integer indices. The rows with unmatched nodes are also dropped.

```{r}
mc3_edges_indexed <- mc3_edges_cleaned %>%
  left_join(node_index_lookup, 
            by = c("from_id" = "id")) %>%
  rename(from = .row_id) %>%
  left_join(node_index_lookup, 
            by = c("to_id" = "id")) %>%
  rename(to = .row_id) %>%
  select(from, to, is_inferred, type) %>%
  filter(!is.na(from) & !is.na(to))  
```

Next, the node list will only include nodes that are used in the edge list and new indices are assigned

```{r}
used_node_indices <- sort(
  unique(c(mc3_edges_indexed$from, 
           mc3_edges_indexed$to)))

mc3_nodes_final <- mc3_nodes_cleaned %>%
  slice(used_node_indices) %>%
  mutate(new_index = row_number())
```

The lookup is rebuild from old index to new index.

```{r}
old_to_new_index <- tibble(
  old_index = used_node_indices,
  new_index = seq_along(
    used_node_indices))
```

Then, we'll update edge indices to match new node table

```{r}
mc3_edges_final <- mc3_edges_indexed %>%
  left_join(old_to_new_index, 
            by = c("from" = "old_index")) %>%
  rename(from_new = new_index) %>%
  left_join(old_to_new_index, 
            by = c("to" = "old_index")) %>%
  rename(to_new = new_index) %>%
  select(from = from_new, to = to_new, 
         is_inferred, type)
```

## 2.5 Build tidygraph Object

Next, we'll build the tidygraph object

```{r}
mc3_graph <- tbl_graph(
  nodes = mc3_nodes_final,
  edges = mc3_edges_final,
  directed = TRUE
)
```

```{r}
str(mc3_graph)
```

## 2.6 Visualize Knowledge Graph

Next, we'll set a seed to ensure reproducibility.

```{r}
set.seed(1234)
```

```{r}
ggraph(mc3_graph, 
       layout = "fr") +
  geom_edge_link(alpha = 0.3, 
                 colour = "gray") +
  geom_node_point(aes(color = `type`), 
                  size = 4) +
  geom_node_text(aes(label = type), 
                 repel = TRUE, 
                 size = 2.5) +
  theme_void()
```

## 3.1 Objective #1: Group Individual and Groups based on the Topics Frequently Brought Up in Their Exchanges

### 3.2 Visualize the Subgraph

A table is generated to see the distinct people, organization and groups whose radio communication Clepper has intercepted.

```{r}
all_nodes <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble()

# Filter by sub_type
persons <- all_nodes %>%
  filter(sub_type == "Person") %>%
  select(name) %>%
  rename(Person = name)

organizations <- all_nodes %>%
  filter(sub_type == "Organization") %>%
  select(name) %>%
  rename(Organization = name)

groups <- all_nodes %>%
  filter(sub_type == "Group") %>%
  select(name) %>%
  rename(Group = name)
max_len <- max(nrow(persons), nrow(organizations), nrow(groups))

table_combined <- tibble(
  Person = c(persons$Person, rep(NA, max_len - nrow(persons))),
  Organization = c(organizations$Organization, rep(NA, max_len - nrow(organizations))),
  Group = c(groups$Group, rep(NA, max_len - nrow(groups)))
)

datatable(table_combined)
```

Next, we'll combine all the people, groups and organizations to check what are the people/organization/groups each individual talk about or talk to:

```{r}
people <- c(persons$Person, groups$Group, organizations$Organization)
persons <- all_nodes %>%
  filter(sub_type == "Person") %>%
  select(name, new_index)
```

### 3.3 1st Individual: Sam

First, we'll look for nodes that are connected to Sam.

```{r}
sam_content <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "1"|to == "1")

used_node_sam <- sam_content %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()
```

Next, we'll check the new_index in the nodes and extract its name, sub type and content

```{r}
connected_node_content_sam <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_sam) %>%
  select(name, sub_type, content)

```

Then we combine every exchanges Sam receives or sends and check for the contents of the exchanges. We can see which organization or group or individual he talks about.

```{r}
all_text_sam <- connected_node_content_sam %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_sam$full_text))

text_clean_sam <- connected_node_content_sam %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

people_clean <- str_to_lower(people) %>%
  str_replace_all("[^a-z\\s]", "") %>%
  str_squish()
```

The result is stored in name_freq_sam, where each occurrence of people/group/organization name will be counted and ordered by the highest occurance of names.

```{r}
name_freq_sam <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_sam$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_sam)
```

Sam seem to talk to the Lookout and Kelly based on the exchanges and he talked a lot about conservation vessels. Additionally, he could be categorized as environmentalist as he and his friends talked a lot about watching animals near or on the Nemo Reef.

### 2nd Individual: Kelly

The same procedure is done for Elise and other individuals to check which people, group or organization they talk about.

```{r}
kelly_content <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "2"|to == "2")

used_node_kelly <- kelly_content %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()
  
connected_node_content_kelly <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_kelly) %>%
  select(name, sub_type, content)

all_text_kelly <- connected_node_content_kelly %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_kelly$full_text))

text_clean_kelly <- connected_node_content_kelly %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_kelly <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_kelly$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_kelly)
```

Elise frequently communicate with Sam about animals and inicidents happening in the Nemo Reef. Similarly to Sam, from the exchanges she had, she can also be categorized as environmentalist especially when she spoek to Sam on the approved permit.

### 3rd Individual: Nadia Conti

```{r}
nadia_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "3"|to == "3")

used_node_nadia <- nadia_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_nadia <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_nadia) %>%
  select(name, sub_type, content)

all_text_nadia <- connected_node_content_nadia %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_nadia$full_text))

text_clean_nadia <- connected_node_content_nadia %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_nadia<- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_nadia$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_nadia)
```

Nadia Conti frequently talk about or communicate to Davis and Elise, Sam, Sailor Shift, the Middleman, the Accountant, Rodriquez, Sailor Shifts Team, V. Miesel Shipping and Conservation Vessels. She can be categorized as Sailor Shift as the team communicates with her on staffing and permit. So it's highly likely that they're working together.

### 4th Individual: Elise

```{r}
elise_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "4"|to == "4")

used_node_elise <- elise_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_elise <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_elise) %>%
  select(name, sub_type, content)

all_text_elise <- connected_node_content_elise %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_elise$full_text))

text_clean_elise <- connected_node_content_elise %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_elise <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_elise$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_elise)
```

Elise frequently communicate to Sam, Nadia, Liam, Neptune, Mako and Ramora. She can be categorized as Sailor Shift as she is involved with Nadia in transferring equipment and the operation.

### 5th Individual: Liam Thorne

```{r}
liam_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "5"|to == "5")

used_node_liam <- liam_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_liam <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_liam) %>%
  select(name, sub_type, content)

all_text_liam <- connected_node_content_liam %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_liam$full_text))

text_clean_liam <- connected_node_content_liam %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_liam <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_liam$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_liam)
```

Liam Thorne seem to be an important person as he works together with with Paackland Harbor and the City Council. However, he can be categorized as Sailor Shifts as he works together with Nadia, possibly with a bribe to ensure that the Harbor Master stays cooperative.

### 6th individual: Samantha Blake

```{r}
samantha_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "6"|to == "6")

used_node_samantha <- samantha_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_samantha <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_samantha) %>%
  select(name, sub_type, content)

all_text_samantha <- connected_node_content_samantha %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_samantha$full_text))

text_clean_samatha <- connected_node_content_samantha %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_samantha <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_samatha$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_samantha)
```

Samantha Blake frequently communicate to Oceanus City Council and Sailor Shift team on music video shooting. However, it seems that she couldn't obtain the permit for it as Nemo Reef is protected areas that they could not shoot there. She can be categorized as Sailor Shift as she works directly to provide them with offiical permit.

### 7th Individual: Davis

```{r}
davis_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "7"|to == "7")

used_node_davis <- davis_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_davis <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_davis) %>%
  select(name, sub_type, content)

all_text_davis <- connected_node_content_davis %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_davis$full_text))

text_clean_davis <- connected_node_content_davis %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_davis <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_davis$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_davis)
```

Davis frequently talk about or communicate to Nadia Conti, Mako, Rodriquez Small Fry and the Middleman. He can be categorized as Sailor Shift as he works with Sailor's team.

### 8th Individual: Rodriquez

```{r}
rodriquez_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "8"|to == "8")

used_node_rodriquez <- rodriquez_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_rodriquez <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_rodriquez) %>%
  select(name, sub_type, content)

all_text_rodriquez <- connected_node_content_rodriquez %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_rodriquez$full_text))

text_clean_rodriquez <- connected_node_content_rodriquez %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_rodriquez <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_rodriquez$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_rodriquez)
```

Rodriquez frequently communicate to Davis, as he is most likely colleagues with Davis and works as logistics. He can be categorized as Sailor Shift as he works directly with Sailor's team for filming location and equipment transfer.

### 9th Individual: Sailor Shifts

```{r}
sailor_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "9"|to == "9")

used_node_sailor <- sailor_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_sailor <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_sailor) %>%
  select(name, sub_type, content)

all_text_sailor <- connected_node_content_sailor %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
all_text_sailor

```

It seems that Sailor, though recorded as an individual, but did not communicate to anyone.

### 10th Individual: Clepper Jensen

```{r}
clepper_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "10"|to == "10")

used_node_clepper <- clepper_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_clepper <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_clepper) %>%
  select(name, sub_type, content)

all_text_clepper <- connected_node_content_clepper %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_clepper$full_text))

text_clean_clepper <- connected_node_content_clepper %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_clepper <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_clepper$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_clepper)
```

Clepper Jensen frequently communicate to Miranda on Rodriquez, Conservation Vessels, Sam, V. Miesel Shipping, sailor Shift, Oceanus City Council and Sailor Shift's team. He is working to expose them and the corruption in Oceanus.

### 11th Individual: Miranda Jordan

```{r}
miranda_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "11"|to == "11")

used_node_miranda <- miranda_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_miranda <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_miranda) %>%
  select(name, sub_type, content)

all_text_miranda <- connected_node_content_miranda %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_miranda$full_text))

text_clean_miranda <- connected_node_content_miranda %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_miranda <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_miranda$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_miranda)
```

Miranda Jordan frequently communicate to Clepper on Rodriquez, Conservation Vessels, Sam, V. Miesel Shipping, Sailor Shift and his/her team and Oceanus City Council. She is working together with Clepper to expose the corruption and manipulation in Oceanus.

### 12th Individual: The Intern

```{r}
intern_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "12"|to == "12")

used_node_intern <- intern_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_intern <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_intern) %>%
  select(name, sub_type, content)

all_text_intern <- connected_node_content_intern %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_intern$full_text))

text_clean_intern <- connected_node_content_intern %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_intern <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_intern$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_intern)
```

The Intern frequently talk about or communicate to the Lookout, Mrs. Money on Conservation Vessels, and V. Miesel Shipping. It's interesting because it doesn't seem that they're working with Sailor's team, but has an operation that they're working on too.

### 13th Individual: The Lookout

```{r}
lookout_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "13"|to == "13")

used_node_lookout <- lookout_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_lookout <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_lookout) %>%
  select(name, sub_type, content)

connected_node_content_lookout <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_lookout) %>%
  select(name, sub_type, content)

all_text_lookout <- connected_node_content_lookout %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_lookout$full_text))

text_clean_lookout <- connected_node_content_lookout %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_lookout <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_lookout$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_lookout)
```

The Intern frequently talk about or communicate to the Lookout, Conservation Vessels, Mrs. Money, Green Guardians and Sam.

### 14th Individual: The Accountant

```{r}
accountant_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "14"|to == "14")

used_node_accountant <- accountant_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_accountant <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_accountant) %>%
  select(name, sub_type, content)

all_text_accountant <- connected_node_content_accountant %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_accountant$full_text))

text_clean_accountant <- connected_node_content_accountant %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_accountant <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_accountant$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_accountant)
```

The Accountant seem to be an important person as he/she spoke with the council and planned to deflect so that they won't get suspicious. He frequently spoke with Boss and the Middleman.

### 15th Individual: Mrs. Money

```{r}
money_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "15"|to == "15")

used_node_money <- money_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_money <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_money) %>%
  select(name, sub_type, content)

all_text_money <- connected_node_content_money %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_money$full_text))

text_clean_money <- connected_node_content_money %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_money <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_money$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_money)
```

Mrs. Money seem to be a person who wants to be anonymous and keep their real identities hidden, as frequently talk about or communicate to the Intern, Boss, the Lookout, the Middleman, people who also use nicknames. Apart from that Mrs. Money seem to frequently communicate with the Lookout, Boss, Mako and the Middleman, they seem to have their own operations on Nemo Reef although it's not very clear what it is.

### 16th Individual: The Middleman

```{r}
middleman_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "16"|to == "16")

used_node_middleman <- middleman_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_middleman <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_middleman) %>%
  select(name, sub_type, content)

all_text_middleman <- connected_node_content_middleman %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) 
cat(as.character(all_text_middleman$full_text))

text_clean_middleman <- connected_node_content_middleman %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_middleman <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_middleman$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_middleman)
```

The Middleman is similar to Mrs. Money, he/she talk about or communicate to Mrs. Money, Boss, the Acountant, people who keep their identities hidden. He/she also talks about or to V. Miesel Shipping, Green Guardians, Sam and Conservation Vessels.

### 17th Inidvidual: Boss

```{r}
boss_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "17"|to == "17")

used_node_boss <- boss_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_boss <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_boss) %>%
  select(name, sub_type, content)

all_text_boss <- connected_node_content_boss %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_boss$full_text))

text_clean_boss <- connected_node_content_boss %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_boss <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_boss$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_boss)
```

Boss frequently talk about or communicate to Mrs. Money and the Middleman for their own operation for tourism ventures.

### 18th Individual: Small Fry

```{r}
fry_contents <- mc3_graph %>%
  activate(edges) %>%
  filter(from == "18"|to == "18")

used_node_fry <- fry_contents %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique()

connected_node_content_fry <- mc3_graph %>%
  activate(nodes) %>%
  as_tibble() %>%
  filter(new_index %in% used_node_fry) %>%
  select(name, sub_type, content)

all_text_fry <- connected_node_content_fry %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " "))
cat(as.character(all_text_fry$full_text))

text_clean_fry <- connected_node_content_fry %>%
  filter(!is.na(content)) %>%
  summarise(full_text = paste(content, collapse = " ")) %>%
  mutate(full_text = str_to_lower(full_text),
         full_text = str_replace_all(full_text, "[^a-z\\s]", ""),
         full_text = str_squish(full_text))

name_freq_fry <- tibble(name = people, clean_name = people_clean) %>%
  rowwise() %>%
  mutate(freq = str_count(text_clean_fry$full_text, fixed(clean_name))) %>%
  arrange(desc(freq))
datatable(name_freq_fry)
```

Small Fry frequently talk about or communicate to Glitter's Team, Davis and Boss. He/she potentially works with Sailor's team but also with Boss.

## 4.1 Objective #2: Is Nadia Conti currently involved in illegal activities ?

### 4.2 Exchanges by or about Nadia Conti

First, we'll look at the communication that mentioned Nadia Conti, Conti and Nadia as the content mentions their identity and to whom they send the message to.

```{r}
comm_about_nadia <- mc3_graph %>%
  activate(nodes) %>%
  filter(str_detect(content, "Nadia Conti| Nadia | Conti ")) %>%
  distinct(content, .keep_all = TRUE)

timeline_nadia_comm <- comm_about_nadia %>%
  as_tibble(active = "nodes")

timeline_nadia_comm <- timeline_nadia_comm %>%
  mutate(
  timestamp = as.POSIXct(timestamp, format = "%Y-%m-%d %H:%M:%S"),
    timestamp_desc = format(timestamp, "%A, %B %d %Y, %H:%M"))

timeline_nadia <- create_tml(df=timeline_nadia_comm, smr="timestamp_desc", dsc="content", smr_bgcol = "#613659", dsc_col = "#613659")
timeline_nadia
```

::: {.callout-note title="Insight from the timeline"}
-   People that are involved: Rodriquez, Mako, Davis, Haackle Harbor, Oceanus City Council, Neptune, Elise, Marlin, Liam, V. Miesel Shipping, V. Miesel HQ, Glitter's Team, The Middleman, Ramora, Harbor Master and even the superstar Sailor Shift
-   Code names that are mentioned: NR-1045, Delta-3 encryption, Berth 14, ALPHA-937, #CR-7844
-   There are couple of striking message mentioned in the exchanges e.g. "keep tourism appearance", "keep operations discreet", "destroy all related documentation", "ensure Harbor Master remains cooperative" or "we need to stay under radar", which sounds sounds suspicious, so we should look deeper into it.
:::

### 4.3 Create timeline for Code Names

First, let's create timeline for code name that is mentioned in the exchanges. In this code, we'll look at the code name **#CR-7844** as it was mentioned frequently.

```{r}
comm_about_cr <- mc3_graph %>%
  activate(nodes) %>%
  filter(if_any(everything(), ~ str_detect(as.character(.), "#CR-7844"))) %>%
  as_tibble(active = "nodes") %>%
  mutate(
    timestamp = as.POSIXct(timestamp, format = "%Y-%m-%d %H:%M:%S"),
    timestamp_desc = format(timestamp, "%A, %B %d %Y, %H:%M")
  )

timeline_table <- comm_about_cr %>%
  filter(!is.na(content))
timeline_cr <-create_tml(df=timeline_table, smr="timestamp_desc", dsc="content", smr_bgcol = "#2E8BC0", dsc_col = "#2E8BC0")
timeline_cr
```

```{r}
comm_about_cr <- comm_about_cr %>%
  filter(is.na(content)) %>%
  select(-timestamp)
datatable(comm_about_cr)
```

::: {.callout-note title="Insight from the timeline and table"}
-   #CR-7844 is a special marine research permit created under V. Miesel to conduct marine research equipment transport into the Nemo Reef, approved by Oceanus City Council. This indicates that V. Miesel is aware and is involved in the operation
-   Similar to the last timeline built, there are frequent messages about transferring equipment through eastern reef routes and bringing security team and extra crew for the operation.
-   The premit was officially used on 11 October for operation. On one of the message from Small Fry to Davis on 12 October 2040, he said that all documentation seem to be in order for tomorrow's operation.
-   They were secretive about the permit and mentioned Delta-3 encryption before the closing of Nemo Reef
:::

Since we found out that that they will start their operation on 13 October, let's filter for all records on that day.

```{r}
comm_13_oct <- mc3_graph %>%
  activate(nodes) %>%
  filter(if_any(everything(), ~ str_detect(timestamp, "2040-10-13"))) %>%
  as_tibble(active = "nodes") %>%
  mutate(
    timestamp = as.POSIXct(timestamp, format = "%Y-%m-%d %H:%M:%S"),
    timestamp_desc = format(timestamp, "%A, %B %d %Y, %H:%M")
  )

timeline_table_comm <- comm_13_oct %>%
  filter(!is.na(content))
timeline_13_oct <-create_tml(df=timeline_table_comm, smr="timestamp_desc", dsc="content", smr_bgcol = "#aa1945", dsc_col = "#aa1945")
timeline_13_oct
```

```{r}
comm_13_oct <- comm_13_oct %>%
  filter(is.na(content)) %>%
  select(-timestamp)
datatable(comm_13_oct)
```

::: {.callout-note title="Insight from the timeline and table"}
-   From the timeline, we can see that the conservation vessel were still monitoring the activities of people with permits. However, what's more surprising is that we found evidence the Oceanus City Council is actually involved in the operation as they told Green Guardians to hold the investigation when there may be potential fuel contamination in the Nemo Reef
-   It seems that there is another operation planned as Nadia asks David to verify the final equipment transport at Nemo Reef.
-   Incidents happening on the day: there was also a music video shooting, equipment transfers and possible fuel contamination on the Nemo Reef
:::

Next, we'll look into the **Delta-3 encryption** since the last exchange about the permit mentioned Delta-3 encryption.So we'll filter exchanges that mentions Delta-3 encryption, Delta or encryption in case they're mentioned separately.

```{r}
comm_encryption <- mc3_graph %>% activate(nodes) %>% filter(str_detect(content, "Delta-3 encryption| encryption | Delta "))

comm_encryption <- comm_encryption %>%
  as_tibble(active = "nodes") %>%  
  select(label, id, timestamp, content, monitoring_type, findings, new_index)

comm_encryption <- comm_encryption %>%
  mutate(
  timestamp = as.POSIXct(timestamp, format = "%Y-%m-%d %H:%M:%S"),
    timestamp_desc = format(timestamp, "%A, %B %d %Y, %H:%M"))

timeline_encryption <- create_tml(df=comm_encryption, smr="timestamp_desc", dsc="content", smr_bgcol = "#1c9b8e", dsc_col = "#1c9b8e")
timeline_encryption
```

::: {.callout-note title="Insight from the timeline"}
-   Delta-3 encryption is used as communication means between everyone that's involved in the operation

-   The operation start executing on the 14 October on 04:00

-   It is confirmed that V. Miesel is involved in the operation as some of its personnel are integrated in the operation
:::

### 4.3 Conclusion from the timeline:

-   Nadia Conti was working together with V. Miesel to create special permits that as intended, she created permit for operations that is illegal

-   There is also bribery done by Nadia as she told Liam that she needed his services to ensure that Harbor Master remains "cooperative" through next week and investigation on Nemo Reef permits were brewing.

Although, there was no evidence apart from her exchanges that she was directly involved in the illegal activities. It seemed like she worked behind the scene in getting the permit and authorizing operations, thus justifying Clepper's suspicions.

## 5. Reference

-   <https://github.com/feddelegrand7/cronologia>
