---
title: "Hands-on Exercise 4d"
author: "Stefanie Felicia"
date: "May, 7 2025"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

# Getting Started

Objective of this exercise:

-   Learn about funnel plots
-   plotting funnel plots by using **funnelPlotR** package
-   plotting static funnel plot by using ggplot2 package
-   plotting interactive funnel plot by using both **plotly R** and **ggplot2**

# Import Libraries

```{r}
pacman::p_load(tidyverse, FunnelPlotR, plotly, knitr)
```

In this exercise, we'll be using the following libraries:

-   plotly

-   tidyverse

-   FunnelPlotR: draws funnel plots using ggplot2

-   knitr: used for dynamic report generation

# Import Data

In this exercise data from <a href="https://riwayat-file-covid-19-dki-jakarta-jakartagis.hub.arcgis.com/">Open Data Covid-19 Provinsi DKI Jakarta portal</a>. We'll be comparing the cumulative COVID-19 cases and death by sub-district (kelurahan).

```{r}
covid19 <- read_csv("COVID-19_DKI_Jakarta.csv") %>%
  mutate_if(is.character, as.factor)
```

::: callout-note
What does the code do:

-   `%>%` passes the tibble to the next function

-   `mutate_if(is.character, as.factor)` line of code checks every column if it's character data type. If it is, it will be stored as integer with labels (or categorical variable)
:::

# FunnelPlotR Package

FunnelPlotR uses ggplot to create funnel plot. The essential arguments are: `numerator` (event of interest), `denominator` (population to be considered) and `group`. The package has the following arguments that can be customized:

-   `limit` : specifies confidence limits for funnel plot (95 or 99)

-   `label_outliers` : label outliers (true or false)

-   `Poisson_limits` : add poisson limit to the plot

-   `OD_adjust` : add overdispersed limit to the plot, which inflates the control limits

-   `xrange` and `yrange` : specify range to display for each axes

-   Aesthetic component

::: {.callout-note title="What is a funnel plot ?"}
-   When we visualize uncertainty, we’ll always be dealing with variation (dealing with multiple occurrences and determine what and how they differ)

-   Two types of variation:

    -   Noise: they are completely random (unknown reasons)

    -   Signal: they are not totally random and can point to underlying causes that can be insightful (to improve)

-   This is where funnel plot comes in! Funnel plot can help us to determine if a data point is noise or signal
:::

# Funnel Plot using FunnelPlotR

The example uses creates funnel plot grouped by sub-district and uses standardized ratio (comapre observed performance agaianst expected) as default plot type.

```{r}
funnel_plot(
  .data = covid19,
  numerator = Positive,
  denominator = Death,
  group = `Sub-district`
)
```

# FunnelPlotR Makeover

```{r}
funnel_plot(
  .data = covid19,
  numerator = Death,
  denominator = Positive,
  group = `Sub-district`,
  data_type = "PR",   
  x_range = c(0, 6500),  
  y_range = c(0, 0.05),   
  label = NA,
  title = "Cumulative COVID-19 Fatality Rate by Cumulative Total Number of COVID-19 Positive Cases",          
  x_label = "Cumulative COVID-19 Positive Cases",
  y_label = "Cumulative Fatality Rate"
)
```

::: {.callout-note title="What does this code do"}
-   `data_type = "PR"` : refers to plotting a proportion ratio plot (used to visualize proportion of percentage of events to compare how different groups perform compared o percentage fo events)

-   `x_range` : sets range of value for x-axis

-   y_range : sets range of value for y-axis

-   `label = NA` : used to remove default label outliers feature
:::

# Funnel Plot using ggplot2

In this example, we want to create funnel plot that plots the cumulative fatality rate by cumulative number of cases.

First, we derive the cumulative death rate and standard error of cumulative death rate.

```{r}
df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  mutate(rate.se = sqrt((rate*(1-rate)) / (Positive))) %>%
  filter(rate > 0)
```

Then we compute the weighted mean.

```{r}
fit.mean <- weighted.mean(df$rate, 1/df$rate.se^2)
```

Next, we compute the lower and upper limit for 95% and 99% CI

```{r}
number.seq <- seq(1, max(df$Positive), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)
```

Now, we can plot the funnel plot

```{r}
p <- ggplot(df, aes(x = Positive, y = rate)) +
  geom_point(aes(label=`Sub-district`), 
             alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            size = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            size = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             size = 0.4, 
             colour = "grey40") +
  coord_cartesian(ylim=c(0,0.05)) +
  annotate("text", x = 1, y = -0.13, label = "95%", size = 3, colour = "grey40") + 
  annotate("text", x = 4.5, y = -0.18, label = "99%", size = 3, colour = "grey40") + 
  ggtitle("Cumulative Fatality Rate by Cumulative Number of COVID-19 Cases") +
  xlab("Cumulative Number of COVID-19 Cases") + 
  ylab("Cumulative Fatality Rate") +
  theme_light() +
  theme(plot.title = element_text(size=12),
        legend.position = c(0.91,0.85), 
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.background = element_rect(colour = "grey60", linetype = "dotted"),
        legend.key.height = unit(0.3, "cm"))
p
```

# Create Interactive Funnel Plot using plotly and ggplot2

We can make the plot above interactive by using ggplotly()

```{r}
fp_ggplotly <- ggplotly(p,
  tooltip = c("label", 
              "x", 
              "y"))
fp_ggplotly
```
