---
title: "Hands-on_Ex10"
author: "Stefanie Felicia"
date: "25 June 2025"
date-modified: "last-modified"
format: html
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Getting Started

Objective of this exercise is to:

-   Create bullet chart using ggplot2

-   Create sparklines using ggplot2

-   Build industry standard dashboard

## Import Libraries

```{r}
pacman::p_load(tidyverse, lubridate, ggthemes, gtExtras, reactable, reactablefmtr)
```

The libraries used in this exercise would be:

-   tidyverse: inclusive of readr, tidyr and dplyr package

-   lubridate: makes it easier to work with dates and times.

-   ggthemes: provide extra themes, geoms and scales from ggplot2

-   gtExtras: extension of the gt package for more customizable gt tables

-   reactable: create interactive data table based on React Table library

-   reactablefmtr: provide various features to streamline and enhance the styling of interactive tables

## The Data

In this hands-on exercise, we'll be using a personal database in Microsoft Access mdb format called Coffee Chain.

## Import Data

We'll use odbcConnectAccess() to import the database query table into R.

```{r}
coffeechain <- read_rds("data/CoffeeChain.rds")
```

## Data Preparation

First, let's aggregate Sales and Budgeted Sales at Product level

```{r}
product <- coffeechain %>%
  group_by(`Product`) %>%
  summarise(`target` = sum(`Budget Sales`),
            `current` = sum(`Sales`)) %>%
  ungroup()
```

## Build Bullet Chart

In this example, we build the bullet chart of all the product to track if they had already "hit the target"

```{r}
ggplot(product, aes(Product, current)) + 
  geom_col(aes(Product, max(target) * 1.01),
           fill="grey85", width=0.85) +
  geom_col(aes(Product, target * 0.75),
           fill="grey60", width=0.85) +
  geom_col(aes(Product, target * 0.5),
           fill="grey50", width=0.85) +
  geom_col(aes(Product, current), 
           width=0.35,
           fill = "black") + 
  geom_errorbar(aes(y = target,
                    x = Product, 
                    ymin = target,
                    ymax= target), 
                width = .4,
                colour = "red",
                size = 1) +
  coord_flip()
```

::: {.callout-note title="What does the code do"}
-   the first 3 `geom_col()` tracks the level of satisfactory of the current product value when it is at that level (e.g bad, okay, satisfactory)
-   the actual value is plotted by the 4th `geom_col()`
-   the target line is plotted by the `geom_errorbar()`
:::

## Plot Sparklines

### Data Preparation

Before we can plot the sparklines, we must prepare the data.

```{r}
sales_report <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  mutate(Month = month(Date)) %>%
  group_by(Month, Product) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup() %>%
  select(Month, Product, Sales)
```

Next, we'll compute the min, max and end of the month sales.

```{r}
mins <- group_by(sales_report, Product) %>% 
  slice(which.min(Sales))
maxs <- group_by(sales_report, Product) %>% 
  slice(which.max(Sales))
ends <- group_by(sales_report, Product) %>% 
  filter(Month == max(Month))
```

We'll also compute the 25 and 75 quantile.

```{r}
quarts <- sales_report %>%
  group_by(Product) %>%
  summarise(quart1 = quantile(Sales, 
                              0.25),
            quart2 = quantile(Sales, 
                              0.75)) %>%
  right_join(sales_report)
```

### Sparkline chart

In this example, we plot the sales per month of each product in the transformed dataset.

```{r}
ggplot(sales_report, aes(x=Month, y=Sales)) + 
  facet_grid(Product ~ ., scales = "free_y") + 
  geom_ribbon(data = quarts, aes(ymin = quart1, max = quart2), 
              fill = 'grey90') +
  geom_line(size=0.3) +
  geom_point(data = mins, col = 'red') +
  geom_point(data = maxs, col = 'blue') +
  geom_text(data = mins, aes(label = Sales), vjust = -1) +
  geom_text(data = maxs, aes(label = Sales), vjust = 2.5) +
  geom_text(data = ends, aes(label = Sales), hjust = 0, nudge_x = 0.5) +
  geom_text(data = ends, aes(label = Product), hjust = 0, nudge_x = 1.0) +
  expand_limits(x = max(sales_report$Month) + 
                  (0.25 * (max(sales_report$Month) - min(sales_report$Month)))) +
  scale_x_continuous(breaks = seq(1, 12, 1)) +
  scale_y_continuous(expand = c(0.1, 0)) +
  theme_tufte(base_size = 3, base_family = "Helvetica") +
  theme(axis.title=element_blank(), axis.text.y = element_blank(), 
        axis.ticks = element_blank(), strip.text = element_blank())
```

::: {.callout-note title="What does the code do"}
-   the `facet_grid()` creates one panel per product (vertically)
-   the `geom_ribbon()` plots shaded area between 1st adn 2nd quantile (which are the min and max respectively)
-   the `geom_line()` plots actual sales over the period of time, while `geom_point()` highlight min value (through red color) and max value (through blue color)
-   the `geom_text()` adds numeric sale values on min and max values and add final month's sales and product name
-   the exapnd_limits()\` expands the x-axis range to leave space for the labels
:::

## Static Information Dashboard Design using gt and gtExtras

### Plot a Simple Bullet Chart

In this example, we plotted a bullet chart using gt_plt_bullet function from the *ft* package, which is definitely simpler than using ggplot2.

```{r}
product %>%
  gt::gt() %>%
  gt_plt_bullet(column = current, 
              target = target, 
              width = 60,
              palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

### Plot Sparklines using gtExtras

Before we can plot the sparklines, we need to prepare the data.

```{r}
report <- coffeechain %>%
  mutate(Year = year(Date)) %>%
  filter(Year == "2013") %>%
  mutate (Month = month(Date, 
                        label = TRUE, 
                        abbr = TRUE)) %>%
  group_by(Product, Month) %>%
  summarise(Sales = sum(Sales)) %>%
  ungroup()
```

gtExtras function require us to pass data.frame with list columns. So in this example, we'll convert the report dataframe into list columns

```{r}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```

Next, we can plot the sales report in form of sparklines.

```{r}
report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop") %>%
   gt() %>%
   gt_plt_sparkline('Monthly Sales',
                    same_limit = FALSE)
```

#### Add statistics for sparklines

First, we can calculate summary statistics of each product (e.g. min, max and average values).

```{r}
report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            ) %>%
  gt() %>%
  fmt_number(columns = 4,
    decimals = 2)
```

Next, we can combine statsitics into the dataframe

```{r}
spark <- report %>%
  group_by(Product) %>%
  summarize('Monthly Sales' = list(Sales), 
            .groups = "drop")
```

```{r}
sales <- report %>% 
  group_by(Product) %>% 
  summarise("Min" = min(Sales, na.rm = T),
            "Max" = max(Sales, na.rm = T),
            "Average" = mean(Sales, na.rm = T)
            )
```

```{r}
sales_data = left_join(sales, spark)
```

#### Plot the updated table with statistics

```{r}
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales',
                   same_limit = FALSE)
```

### Combining Bullet Chart and Sparklines

We can combine the bullet chart with sparklines

```{r}
bullet <- coffeechain %>%
  filter(Date >= "2013-01-01") %>%
  group_by(`Product`) %>%
  summarise(`Target` = sum(`Budget Sales`),
            `Actual` = sum(`Sales`)) %>%
  ungroup() 
```

```{r}
sales_data = sales_data %>%
  left_join(bullet)
```

```{r}
sales_data %>%
  gt() %>%
  gt_plt_sparkline('Monthly Sales') %>%
  gt_plt_bullet(column = Actual, 
                target = Target, 
                width = 28,
                palette = c("lightblue", 
                          "black")) %>%
  gt_theme_538()
```

## Interactive Information Dashboard Design using reactable and reactablefmtr

To build an interactive sparklines, we need to install datui package and load it

```{r, eval=FALSE}
remotes::install_github("timelyportfolio/dataui")
```

```{r}
library(dataui)
```

#### Plot Interactive Sparklines

Similar to the gtExtras package, we need to convert the dataframe to a list column.

```{r}
report <- report %>%
  group_by(Product) %>%
  summarize(`Monthly Sales` = list(Sales))
```

react_sparkline will be used to plot the sparklines

```{r}
reactable(
  report,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

Currently, the sparklines are seperated into 2 page and we want to display it all in one page by changing `defaultPageSize`.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(report)
    )
  )
)
```

Next, we can also improve our current chart by adding points to indicate the min and max values of each product by using the `highlight_points` argument.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        labels = c("first", "last")
        )
    )
  )
)
```

We can also improve our current graph by adding a reference line so that readers can understand the context and interpret the data more accurately (`statline` argument).

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        statline = "mean"
        )
    )
  )
)
```

We can also use bandline instead of reference line by using the `bandline` argument.

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkline(
        report,
        highlight_points = highlight_points(
          min = "red", max = "blue"),
        line_width = 1,
        bandline = "innerquartiles",
        bandline_color = "green"
        )
    )
  )
)
```

Additionally, we can also change the display from sparkline into sparkbar

```{r}
reactable(
  report,
  defaultPageSize = 13,
  columns = list(
    Product = colDef(maxWidth = 200),
    `Monthly Sales` = colDef(
      cell = react_sparkbar(
        report,
        highlight_bars = highlight_bars(
          min = "red", max = "blue"),
        bandline = "innerquartiles",
        statline = "mean")
    )
  )
)
```
