---
title: "Hands-on Exercise 8c"
author: "Stefanie Felicia"
date: "11 June 2025"
date-modified: "last-modified"
format: html
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Getting Started

Objective of this exercise:

-   Import geospatial data in rds format into R
-   Create cartographic quality choropleth maps using appropriate map function
-   Create rate map
-   Create percentile map
-   Create boxmap

## Import Libraries

```{r}
pacman::p_load(tmap, tidyverse, sf)
```

The libraries used in this exercise would be:

-   sf: support for simple feature access, standardized way to encode and analyze spatial vector data

-   tmap: used to draw thematic maps

-   tidyverse: inclusive of readr. tidyr and dplyr package

## The Data

For this exercise, we'll be using a dataset called NGA_wp.rds which is a polygon feature data.frame. It contains information on water point fo Nigeria at the LGA level

```{r}
NGA_wp <- read_rds("data/rds/NGA_wp.rds")
```

## Create Choropleth Map

In this example, we'll be visualizing distribution of non-functional water point

```{r}
p1 <- tm_shape(NGA_wp) + 
  tm_polygons(fill = "total_wp",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1, 
             fill_alpha = 1) + tm_title("Distribution of Functional Water by LGAs")
```

```{r}
p2 <- tm_shape(NGA_wp) + 
  tm_polygons(fill = "total_wp",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1, 
             fill_alpha = 1) + tm_title("Distribution of Total Water by LGAs")
```

```{r}
tmap_arrange(p1, p2, nrow = 1)
```

## Choropleth Map for Rates
