---
title: "Hands-on Exercise 9a"
author: "Stefanie Felicia"
date: "18 June 2025"
date-modified: "last-modified"
format: html
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Getting Started

Objective of this exercise:

-   Derive 3 different measures using mutate() function

-   Build static ternary plot using ggtern() from ggtern package

-   Build interactive ternary lot using plot-ly() from Plotly R

## Import Libraries

```{r}
pacman::p_load(plotly, ggtern, tidyverse)
```

The libraries used in this exercise would be:

-   ggtern: plots ternary diagrams using ggplot2 geometries

-   Plotly R: creates interactive graphs

-   tidyverse: inclusive of readr. tidyr and dplyr package

## The Data

We'll be using te Singapore Residents by Planning Area Subzone, Age Group, Sex and Type of Dwelling, June 2000-2018 from the \<a href ="<https://www.singstat.gov.sg/find-data/search-by-theme/population/geographic-distribution/latest-data>\>Singstat</a> website.

## Import Data

```{r}
pop_data <- read_csv("data/respopagsex2000to2018_tidy.csv") 
```

### Data Preparation

Next, we'll derive 3 new variables called young, active and active.

```{r}
#Deriving the young, economy active and old measures
agpop_mutated <- pop_data %>%
  mutate(`Year` = as.character(Year))%>%
  spread(AG, Population) %>%
  mutate(YOUNG = rowSums(.[4:8]))%>%
  mutate(ACTIVE = rowSums(.[9:16]))  %>%
  mutate(OLD = rowSums(.[17:21])) %>%
  mutate(TOTAL = rowSums(.[22:24])) %>%
  filter(Year == 2018)%>%
  filter(TOTAL > 0)
```

## Plotting Ternary Diagram with R

Next, we'll build a static ternary plot using ggtern() to plot the variables we've created in the data preparation steps.

```{r}
ggtern(data=agpop_mutated,aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point()
```

But we can add colors and title to make it clearer what we're plotting by adding `theme_rgbw()` to differentiate between variables and `labs(title...)`.

```{r}
ggtern(data=agpop_mutated, aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point() +
  labs(title="Population structure, 2015") +
  theme_rgbw()
```

## Plotting Interactive Ternary Plot

We'll plot interactive ternary plot using plot-ly()

```{r}
# reusable function for creating annotation object
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

# reusable function for axis formatting
axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young"), 
  baxis = axis("Active"), 
  caxis = axis("Old")
)

# Initiating a plotly visualization 
plot_ly(
  agpop_mutated, 
  a = ~YOUNG, 
  b = ~ACTIVE, 
  c = ~OLD, 
  color = I("black"), 
  type = "scatterternary"
) %>%
  layout(
    annotations = label("Ternary Markers"), 
    ternary = ternaryAxes
  )
```
