---
title: "Hands-on_Ex09e"
author: "Stefanie Felicia"
date: "20 June 2025"
date-modified: "last-modified"
format: html
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Getting Started

Objective of this exercise is:

-   Learn how to manipulate transaction data into treemap structure by using dplyr package

-   Plot static treemap using treemap package

-   How to design interactive treemap with d3treeR package

## Import Libraries

```{r}
pacman::p_load(treemap, treemapify, tidyverse) 
```

The libraries used in this exercise would be:

-   treemap: space-filling visualization of hierarchical strudctures

-   treemapify: draw treemap using ggplot2

-   tidyverse: inclusive of readr. tidyr and dplyr package

## The Data

In this exercsie, we'll be using the REALIS2018.csv data which comprise of private property transaction record in 2018.

## Import Data

```{r}
realis2018 <- read_csv("data/realis2018.csv")
```

## Data Wrangling

Currently, the dataset is in transaction record form, which need manipulation and preparation to plot a treemap. The following steps will be taken to prepare the dataset:

-   Group transaction records by *Project Name, Planning Region, Planning Area, Property Type* and *Type of Sale*

-   Compute *Total Unit Sold*, *Total Area*, *Median Unit Price* and *Median Transacted Price* by applying appropriate summary statistics on *No. of Units*, *Area (sqm)*, *Unit Price (\$ psm)* and *Transacted Price (\$)*

We'll be using two functions to all the steps:

-   group_by() breaks dataframe into specified group of rows

-   summarize() to derive summary of each group

### Grouped summaries without the Pipe

```{r}
realis2018_grouped <- group_by(realis2018, `Project Name`,
                               `Planning Region`, `Planning Area`, 
                               `Property Type`, `Type of Sale`)
realis2018_summarised <- summarise(realis2018_grouped, 
                          `Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE),
                          `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
                          `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE), 
                          `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```

### Grouped Summaries with the Pipe

```{r}
realis2018_summarised <- realis2018 %>% 
  group_by(`Project Name`,`Planning Region`, 
           `Planning Area`, `Property Type`, 
           `Type of Sale`) %>%
  summarise(`Total Unit Sold` = sum(`No. of Units`, na.rm = TRUE), 
            `Total Area` = sum(`Area (sqm)`, na.rm = TRUE),
            `Median Unit Price ($ psm)` = median(`Unit Price ($ psm)`, na.rm = TRUE),
            `Median Transacted Price` = median(`Transacted Price ($)`, na.rm = TRUE))
```

::: {.callout-note title="What's the difference between both code"}
-   The pipe (%\>%) chains together the multiple methods calls, which in this case chains together the group_by() function and summarise() function that they don't need to be stored in separate variables.
:::

## Design Treemap using treemap Package

The treemap() function offers 43 different arguments, which can be found <a href="https://www.rdocumentation.org/packages/treemap/versions/2.4-4/topics/treemap">here</a>. However, we'll only be exploring the major arguments to design elegant and truthful treemaps

### Design static treemap

In this example, we'll be plotting a treemap showing the distribution of median unit prices and total unit sold of resale condominium by geographic hierarchy in 2017.

So. we'll select records that are resale condominium property type.

```{r}
realis2018_selected <- realis2018_summarised %>%
  filter(`Property Type` == "Condominium", `Type of Sale` == "Resale")
```

Next, we'll use the following arguments to draw the treemap:

-   index: The index vector must consist of at least 2 column names or else the plot will be empty. If multiple columns are provided, the first name is the highest aggregation level and then the second name, to the third, and so on.

-   vSize: column that determines the rectangle area size, this means that column must not contain negative values

-   vColor: column that determine the rectangle's color, used in combination with type argument.

```{r}
treemap(realis2018_selected,
            index=c("Planning Region", "Planning Area", "Project Name"),
            vSize="Total Unit Sold",
            vColor="Median Unit Price ($ psm)",
            title="Resale Condominium by Planning Region and Area, 2017",
            title.legend = "Median Unit Price (S$ per sq. m)"
            )
```

### Working with vColor and type arguments

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type = "value",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

As can be seen, the type argument changes the color scheme and arrangement of the treemap.

### Colors in the treemap package

Arguments that determine the mapping to color palattes are: mapping and palette. The difefrence between "value" and "manual" is the default value for mapping. The "value" treemap considers palette to be a diverging color palette (e.g. RdYIBu) and map it in a way that 0 corresponds to the middle color (usually white or yellow), -max(abs(values)) to the left-end color and max(abs(values)) to the left-end color, max(values) to the right-end color and mean(range(values)) to the middle color.

#### The "value" type treemap

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="value",
        palette="RdYlBu", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

::: {.callout-note title="Points on the code"}
-   The color palette used RdYIBu palette, but there's no red rectangle in the treemap because the median unit prices are positive
-   The median Unit Price is from 5000 - 45000 in the legends because the range argument by default c(min(values,max(values)) with rounding.
:::

#### The "manual" type treemap

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="RdYlBu", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

The color schemes of the two type of treemap is completely different even though they use the same palette. this is because "manual" type doesn't interpret values like "values" does. The value range is mapped linearly to the color palette. In "values", the mapping is (min(values, mean(range(values)), max(values)). In this example, we don't need to use diverging color palette because the values are all positive (this works too if values are all negative).

We can change the color palette to blues, which only uses single color palette.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

### Treemap Layout

There are two popular treemap layout in treemap(): "squarified" and "pivotSize" (default).

The squarified treemap algorithm produces godo aspect ratios, but ignores the sorting order of the rectangles (sortID). Th ordered treemap, pivot-by-size, algorithm takes sorting order (sortID) into account, whole aspect ratio is still acceptable.

#### Working with algorithm argument

In this example, we'll plot the squarified treemap by changing the algorithm argument.

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "squarified",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

#### Using sortID

When the pivotSize algorithm is used, sortID argument can be used to determine order (the rectangles are placed from top left to bottom right).

```{r}
treemap(realis2018_selected,
        index=c("Planning Region", "Planning Area", "Project Name"),
        vSize="Total Unit Sold",
        vColor="Median Unit Price ($ psm)",
        type="manual",
        palette="Blues", 
        algorithm = "pivotSize",
        sortID = "Median Transacted Price",
        title="Resale Condominium by Planning Region and Area, 2017",
        title.legend = "Median Unit Price (S$ per sq. m)"
        )
```

## Design Treemap using treemapify

treemapify uses ggplot2 to plot the treemap. In this example, we can create basic treemap like so.

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`),
       layout = "scol",
       start = "bottomleft") + 
  geom_treemap() +
  scale_fill_gradient(low = "light blue", high = "blue")
```

Next, we can define the hierarchy (group them by planning region)

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`),
       start = "topleft") + 
  geom_treemap()
```

If we also group by planning area, the layout will look vastly different.

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap()
```

Next, we want to add boundary line

```{r}
ggplot(data=realis2018_selected, 
       aes(area = `Total Unit Sold`,
           fill = `Median Unit Price ($ psm)`,
           subgroup = `Planning Region`,
           subgroup2 = `Planning Area`)) + 
  geom_treemap() +
  geom_treemap_subgroup2_border(colour = "gray40",
                                size = 2) +
  geom_treemap_subgroup_border(colour = "gray20")
```

Designing Interactive Treemap using d3treeR

Make sure that you already have devtools package installed. If not download using this code

```{r, eval=false}
install.packages("devtools")
```
