---
title: "Hands-on Exercise 9c"
author: "Stefanie Felicia"
date: "19 June 2025"
date-modified: "last-modified"
format: html
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Getting Started

Objective of this exercise is to plot static and interactive heatmap to visualize and analyze multivariate data, as heatmaps are useful to reveal patterns, display whether any variables are similar to each other and if there's any correlations that exist in-between them.

## Import Libraries

```{r}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)
```

The libraries used in this exercise would be:

-   seriation: infrastructure for ordering objects with an implementation of several seriazation or sequencing or ordination techniques to reorder matrices, dendograms and dissimilarity matrices.

-   dendextend: offer set of functions for extending dendrogram object in R, to visualize and compare trees of hierarchical clustering

-   heatmaply: create interactive heatmaps that can be standalone HTML file

-   tidyverse: inclusive of readr, tidyr and dplyr package

## The Data

In this hands-on, we'll be using the World Happiness 2018 report dataset.

## Import Data

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

## Data Preparation

We want to change the rows by country name instead of row number

```{r}
row.names(wh) <- wh$Country
```

Next, we want to change the dataframe into data matrix to create the heatmap.

```{r}
wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

## Static Heatmap Plot inR

There re a few R paccakages that can be used to plot static heatmap:

-   heatmap(): build a simple heatmap

-   heatmap.2(): is enhanced heatmap (more advanced than heatmap())

-   heatmap() from pheatmap package: has monroe control over the aesthetics parts

-   ComplexHeatmap: draws, annotate and arrange complex heatmap

-   Superheat package: exploring complex dataset using heatmap (even big data with different data types

-   Heatmaply: package to build interactive heatmap (can be shared as HTML standalone file)

## Visualize heatmap using heatmap()

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      Rowv=NA, Colv=NA)
```

::: callout-note
-   By default, heatmap() plots a cluster heatmap. The arguments Rowv=NA and Colv=NA are used to switch off the option of plotting the row and column of dendogram
:::

If we want to change to visualize cluster heatmap we can use the default code without any changes like below

```{r}
wh_heatmap <- heatmap(wh_matrix)
```

::: callout-note
-   We can see that the order of the dendogram is different from the heatmap we create earlier because heatmap reorder using clusterization (calculate the distance between each pair of rows and columns and try to order them by similarity).
-   The red cell denotes high values, whereas yellow cells represent low value, this can be quite confusing especially when most of the heatmap and dendogram shows the yellow cells. We need to normalize the matrix by using `scale` argument, which can be applied to rows or columns.
:::

In this example, we'll normalize the matrix (column wise)

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

::: {.callout-note title="What does the code do"}
-   `margin` argument is used to ensure the entire x-axis labels are displayed completely
-   `cexRow` and `cexCol` argument are used to define font size used for y-axis and x-axis
:::

## Create Interactive Heatmap

In this example, we'll create interactive heatmap using <a href="https://cran.r-project.org/web/packages/heatmaply/vignettes/heatmaply.html">heatmaply()</a>.

```{r}
heatmaply(mtcars)
```

Now, let's create another heatmap using the basic syntax to create n interactive heatmap

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)])
```

### Data Transformation

In order to ensure that all variable have comparable values (same range throughout all variables so that the comparison is fair between all variable), so data transformation are applied before clustering.

In heatmaply(), there are 3 main data transformations supported: scale, normalise and percentile.

#### Scaling Method

-   This method is used when all variables come form some normal distribution. Scaling would bring them all close to normal standard distribution by subtracting the mean and divide by standard deviation.
-   Each value would reflect the distance from the means in its units of standard deviation
-   The scale argument in heatmaply() supports column and row scaling

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          scale = "column")
```

#### Normalizing Method

This method is used when variables in data come from non-normal distributions, the normalize function can be used to bring data from 0 to 1 scale by subtracting the minimum and dividing by the maximum of all observations. This implies that the shape of each variable's distribution while making them comparable on the same scale

In this example, the normalize method is performed on the input data set like below

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]))
```

#### Percentizing Method

-   It works by ranking the variables and divide them by the maximal rank (instead of keeping the rank values). By using the ecdf of the variables on their own values, each value is brought to its empirical percentile
-   Benefit of the percentile function is that each value has relatively clear interpretation and it is the percent of observations that got that value or below it. In this example, the percentize method is performed on the input data like below.

```{r}
heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]))
```

### Clustering Algorithm

heatmaply supports different hierarchical clustering algorithm. The main arguments are:

-   distfun: used to compute the distance (dissimilarity) between both rows and columns. There are "pearson", "spearman" and "kendall" options to be used for correlation-based clustering. It uses as.dist(1-cor(t(x))) as the distance metric using the specified correlation option.

-   hclustfun: used to compute hierarchical clustering when Rowv or Colv are not dendograms

-   dist_method: its default is NULL and results are in "euclidean". It can accept alternative character strings indicating the method to be passed to distfun. By default distfun is "dist", so it can be one of "euclidean", "maximum", "manhattan", "canberra" or "minkowski"

-   hclust_method: its default is NULL and results in "complete" method. It can accept alternative character strings indicating the method to be passed to hclust_fun. By default hclustfun is hclust, so it can be one of "ward.D", "ward.D2", "single", "complete", "average" (UPGMA), "mcquitty" (WPGMA), "median" (WPGMC) or "centroid" (UPGMC)

Clustering model can be produced manually or statistically.

#### Manual Approach

In this example, the heatmap is plotted using hierarchical clustering algorithm with euclidean distance and ward.D method

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```

#### Statistical Approach

We can use dend_expend() to find the best clustering method.

```{r}
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]
```

Then use find_k() to find best number of cluster.

```{r}
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

We can see that 3 would be the best number of cluster and plot the heatmap

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```

### Seriation

Seriation package is used to find optimal ordering of rows and columns since hierarchical clustering doesn't place rows in a definite order, it only give you possible ordering (not which one would lead to the best clearer-looking heatmap). It optimizes by minimizing the Hamiltonian path length that is restricted by dendogram structure.

Seriation algorithm:

-   Optimal Leaf Ordering (OLO), which start with the output of agglomerative clustering algorithm and produce unique ordering. It flips the various branches of the dendogram around to minimize the sum of dissimilarities between adjacent leaves. The algorithm optimizes in O(n\^4). The example below applies OLO algorithm:

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```

-   Gruvaeus and Wainer (GW): similar to OLO algorithm, but uses potentially faster heuristic

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")
```

-   mean: gives default results from heatmap function from other packages (gplots::heatmap.2)

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "mean")
```

-   none: dendogram without any rotation based on data matrix

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none")
```

### Working with Color Palette

Color palette in heatmaply by default uses viridis. But there are more color palette we can use. In this example, we use the Blues color palette of rColorBrewer

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Blues)
```

Extra information

Heatmaply also provide many plotting features to ensure cartographic quality is ensured. In this example, k_row is used to produced 5 rows, whereas margins is used to change top margin and row margin. fontsize_row and fontsize_col are used to change font size for row and columns, while main is used to write the main title fo the plot and xlab and ylab are used to write x-xis adn y-axis label

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```
