---
title: "Hands-on Exercise 9d"
author: "Stefanie Felicia"
date: "22 June 2025"
date-modified: "last-modified"
format: html
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## Getting Started

Objective of this exercise is:

-   Plot static parallel coordinate plot using ggparcoord() of GGally package

-   Plot interactive parallel coordinate plot using ParallelPlot package

## Import Libraries

```{r}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

The libraries used in this exercise would be:

-   GGally: plotting system extended from ggplot2

-   parallelPlot: create parallel coordinates ploy

-   tidyverse: inclusive of readr. tidyr and dplyr package

## The Data

In this hands-on, we'll be usig the World Hapiness 2018 datataset.

## Import Data

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

## Plotting Static Parallel Coordinates Plot

The below shows the basic syntax needed to build parallel coordinate plot using ggparcoord().

```{r}
ggparcoord(data = wh, 
           columns = c(7:12))
```

The data argument is used to map the data object, whereas columns argument is used to select the columns to prepare the parallel coordinate plot.

## Plotting Parallel Coordinates with boxplot

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")
```

::: {.callout-note title="What does the code do"}
-   `groupColumn` argument is used to group the observations (parallel lines) by using single variable (region) and color the parallel coordinates lines by region name
-   `scale` argument is used to scale the variables in the parallel coordinate plot by using `uniminmax` method. The method univariately scale each variable so the min of the variable is 0 and the max is 1
-   `alphaLines` argument is used to reduce the intensity of the line color to 0.2. The permissible value range is from 0 to 1.
-   `boxplot` argument is used to turn on the boxplot by using logical TRUE, which the default value is FALSE.
-   `title` argument is used to provide the parallel coordinates boxplot a title
:::

Parallel Coordinates with Facet We can combine some of the ggplot2 function when plotting parallel coordinate plot. In thisexample, we use `facet_wrap()` to plot 10 small multiple coordinate plots. Each plot represent one region.

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region)
```

### Rotating x-axis Text Label 

As can be seen from the plot above, some of the variable names overlap on x-axis. So we can rotate the them by 30 degrees. This can be done using `theme()` (axis.text.x)

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))
```

### Adjust Rotated x-axis Text Label

The plot above still show overlapping if x-axis text label with the plot. We can adjust text location by using `hjust` argument to theme's text element with `element_text()` on top of the existing `axis.text.x` argument.

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

## Plotting Interactive Parallel Coordinate Plot

The code shows the basic syntax to plot interactive parallel coordinate plot with parallelPlot().

```{r}
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh,
             width = 320,
             height = 250)
```

### Rotate Axis Label

As can be seen from the previous plot, some of the axis label are too long which we can resolve by adding `rotateTitle=TRUE`

```{r}
parallelPlot(wh,
             rotateTitle = TRUE)
```

### Change Color Scheme

By default the color scheme is blue, but we can change that by using `continuousCS` argument.

```{r}
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```

### Parallel Coordinate Plot with Histogram

In this example, we plot histogram alongside the axis of each variable using `histoVisibility` argument.

```{r}
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```
